##################################################################################
#                                                                                #
#  TITLE: The 'WXYZ' model - probabilistic                                       #
#                                                                                #
#  AUTHOR: Ian Andrew Cromwell, PhD                                              #
#                                                                                #
#  NOTE: This code is a modified verion of a program written by the              #
#        DARTH Working Group. Full attribution follows.                          #
#                                                                                #
#  PURPOSE: This code is a deterministic version of the Markov model that is     #
#            described on the 'Healthy Uncertainty' R Guide:                     #
#                 https://healthyuncertainty.github.io/RGuide/Chapter5/          #
#            It has been updated to reflect the DARTH method.                    #
#                                                                                #
##################################################################################

Developed by the Decision Analysis in R for Technologies in Health (DARTH) workgroup:

Fernando Alarid-Escudero, PhD (1) 

Eva A. Enns, MS, PhD (2)	

M.G. Myriam Hunink, MD, PhD (3,4)

Hawre J. Jalal, MD, PhD (5) 

Eline M. Krijkamp, MSc (3)	

Petros Pechlivanoglou, PhD (6,7)

Alan Yang, MSc (7)

In collaboration of: 		

1. Drug Policy Program, Center for Research and Teaching in Economics (CIDE) - CONACyT, 
   Aguascalientes, Mexico
2. University of Minnesota School of Public Health, Minneapolis, MN, USA
3. Erasmus MC, Rotterdam, The Netherlands
4. Harvard T.H. Chan School of Public Health, Boston, USA
5. University of Pittsburgh Graduate School of Public Health, Pittsburgh, PA, USA
6. University of Toronto, Toronto ON, Canada
7. The Hospital for Sick Children, Toronto ON, Canada

Please cite our publications when using this code:
 
- Jalal H, Pechlivanoglou P, Krijkamp E, Alarid-Escudero F, Enns E, Hunink MG. 
An Overview of R in Health Decision Sciences. Med Decis Making. 2017; 37(3): 735-746. 
https://journals.sagepub.com/doi/abs/10.1177/0272989X16686559
 
- Krijkamp EM, Alarid-Escudero F, Enns EA, Jalal HJ, Hunink MGM, Pechlivanoglou P. 
Microsimulation modeling for health decision sciences using R: A tutorial. 
Med Decis Making. 2018;38(3):400–22. 
https://journals.sagepub.com/doi/abs/10.1177/0272989X18754513
 
- Krijkamp EM, Alarid-Escudero F, Enns E, Pechlivanoglou P, Hunink MM, Jalal H. 
A Multidimensional Array Representation of State-Transition Model Dynamics. 
Med Decis Making. Online First https://doi.org/10.1177/0272989X19893973

Copyright 2017, THE HOSPITAL FOR SICK CHILDREN AND THE COLLABORATING INSTITUTIONS. 
All rights reserved in Canada, the United States and worldwide. Copyright, 
trademarks, trade names and any and all associated intellectual property are 
exclusively owned by THE HOSPITAL FOR Sick CHILDREN and the collaborating 
institutions. These materials may be used, reproduced, modified, distributed 
and adapted with proper attribution.


# 01 Load packages

```{r, warning = F, message = F}
if (!require('pacman')) install.packages('pacman'); library(pacman) # use this package to conveniently install other packages
# load (install if required) packages from CRAN
p_load("here", "dplyr", "devtools", "scales", "ellipse", "ggplot2", "lazyeval", "igraph", "truncnorm", "ggraph", "reshape2", "knitr", "stringr", "diagram", "BCEA", "readxl")                                               
# load (install if required) packages from GitHub
# install_github("DARTH-git/dampack", force = TRUE) Uncomment if there is a newer version
# install_github("DARTH-git/dectree", force = TRUE) Uncomment if there is a newer version
p_load_gh("DARTH-git/dampack", "DARTH-git/dectree")
```

# 02 Load functions

```{r}
source('functions_WXYZ.R')
source("ImportVars.R")
```

# 03 Input model parameters

```{r}
# Set number of probabilistic simulations
n_sim <- 1000

# Set number of cycles
n_t     <- 20

# Define societal WTP
lambda <- 10000

# Discounting factor for effects and costs
d_e <- 0.015 
d_c <- 0.015
```

```{r}
# Strategy names
v_names_str <- c("Standard_of_Care", "New_Treatment") 

# Number of strategies
n_str <- length(v_names_str)

# Markov model parameters
v_n     <- c("W", "X", "Ytransition", "Y", "Ztransition", "Z")  # health state names  
n_states     <- length(v_n)         # number of health states

# calculate discount weights for costs for each cycle based on discount rate d_c
v_dwc <- 1 / (1 + d_e) ^ (0:n_t) 
# calculate discount weights for effectiveness for each cycle based on discount rate d_e
v_dwe <- 1 / (1 + d_c) ^ (0:n_t) 
```

```{r}
# Run the batch importer
Inputs <- read_excel("Model Inputs.xls") #The Excel file name
Inputs <- subset(Inputs, Value >-1) #Remove blank rows
varlist <- ImportVars(Inputs, n_sim)
# 'n_sim' probabilistic values are stored in a dataframe called 'df_values'
df_values <- varlist$df_psa_input

# calculate discount weights for costs for each cycle based on discount rate d_c
v_dwc <- 1 / (1 + d_e) ^ (0:n_t) 
# calculate discount weights for effectiveness for each cycle based on discount rate d_e
v_dwe <- 1 / (1 + d_c) ^ (0:n_t) 

```

# 04 Run the Markov Model
START PROBABILISTIC LOOP
Within each loop:
1 - create the Markov trace matrix
2 - load a set of probabilistic parameter estimates with 'get_values'
3 - populate the transition probability matrix with 'make_matrix'
4 - run the Markov model
5 - calculate costs and QALYs with 'make_cea'
6 - record outputs

```{r}
# Arrays to hold the Markov trace outputs
  mtraceout_trt <- mtraceout_notrt <- array(0, dim = c(n_t+1, n_states, n_sim))

# Dataframes to hold the CEA results
  df_c <- df_e <- as.data.frame(matrix(0,
                               nrow = n_sim,
                               ncol = n_str))
  colnames(df_c) <- colnames(df_e) <- v_names_str

# Calculate 'n_sim' probabilistic simulations of the Markov output
for (j in 1:n_sim){
  # 1 - Create the markov trace matrix
    m_M_notrt <- m_M_trt <- matrix(NA, 
                                    nrow     = n_t + 1, ncol = n_states,
                                    dimnames = list(paste("cycle", 0:n_t, sep = " "), v_n))
  
  # 2 - Load a set of probabilistic parameter estimates with 'get_values'
    # Take the values from row 'j' of the list of parameter values
    list_values <- get_values(df_values[j,])
  
  # 3 - populate the transition matrix with 'make_matrix'
    # calculate the transition matrices based on the values in 'list_values'
    list_tmatrix <- make_matrix(list_values)
  
  # 4 - run the Markov model
    # Populate the first row of the trace matrix
    m_M_notrt[1, ] <- m_M_trt[1, ] <- c(list_values$p_W, list_values$p_X, 0, 0, 0, 0)
    # Loop through the number of cycles
    for (t in 1:n_t){
      # Estimate the Markov trace for each cycle 't'
      m_M_notrt[t + 1, ] <- t(m_M_notrt[t, ]) %*% list_tmatrix$m_P_notrt
      m_M_trt[t + 1, ]    <- t(m_M_trt[t, ])  %*% list_tmatrix$m_P_trt      
    }
  
  # 5 - calculate costs and QALYs with 'make_cea'
    cea_out <- make_cea(list_values, m_M_notrt, m_M_trt, v_dwe, v_dwc)
    
  # 6 - record outputs
    mtraceout_notrt[,,j] <- m_M_notrt
    mtraceout_trt[,,j]   <- m_M_trt
    df_c[j,]             <- cea_out$Cost
    df_e[j,]             <- cea_out$Effect
  
  } # End loop
# Restart until there are 'n_sim' results in each output list
```

# 05 Compute Cost-Effectiveness Results with 'dampack'
```{r}
# Create a PSA object
l_psa <- make_psa_obj(cost          = df_c, 
                      effectiveness = df_e, 
                      parameters    = df_values, 
                      strategies    = v_names_str)

# Mean and incremental costs/QALYs, mean ICER
psa_sum <- summary(l_psa, calc_sds = TRUE)  # Estimate the mean costs, QALYs  
df_cea <- calculate_icers(cost       = psa_sum$meanCost,
                          effect     = psa_sum$meanEffect,
                          strategies = psa_sum$Strategy)

# Calculate percent cost-effectiveness of each strategy
v_wtp        <- seq(0,2*lambda, by = lambda/20) # Vector of WTP values from 0 to 2*lambda
ceac_obj     <- ceac(psa = l_psa, wtp = v_wtp)
summary_ceac <- summary(ceac_obj)

# Calculate probability that new treatment is cost-effective at lambda
n_midwtp       <- length(v_wtp)+1            
n_trtPercentCE <- 100*ceac_obj$Proportion[n_midwtp]  

# View results
df_cea                                      
ceac_obj
n_trtPercentCE

# Plot the CEAC
plot(ceac_obj, 
     frontier = TRUE, 
     points = TRUE)
```
