##################################################################################
#                                                                                #
#  TITLE: The 'WXYZ' model                                                       #
#                                                                                #
#  AUTHOR: Ian Andrew Cromwell, PhD                                              #
#                                                                                #
#  NOTE: This code is a lightly modified verion of a program written by the      #
#        DARTH Working Group, who deserve the bulk of authorship credit. Full    #
#        attribution follows.                                                    #
#                                                                                #
#  PURPOSE: This code is a deterministic version of the Markov model that is     #
#            described on the 'Healthy Uncertainty' R Guide:                     #
#                 https://healthyuncertainty.github.io/RGuide/Chapter5/          #
#            It has been updated to reflect the DARTH method.                    #
#                                                                                #
##################################################################################

Developed by the Decision Analysis in R for Technologies in Health (DARTH) workgroup:

Fernando Alarid-Escudero, PhD (1) 

Eva A. Enns, MS, PhD (2)	

M.G. Myriam Hunink, MD, PhD (3,4)

Hawre J. Jalal, MD, PhD (5) 

Eline M. Krijkamp, MSc (3)	

Petros Pechlivanoglou, PhD (6,7)

Alan Yang, MSc (7)

In collaboration of: 		

1. Drug Policy Program, Center for Research and Teaching in Economics (CIDE) - CONACyT, 
   Aguascalientes, Mexico
2. University of Minnesota School of Public Health, Minneapolis, MN, USA
3. Erasmus MC, Rotterdam, The Netherlands
4. Harvard T.H. Chan School of Public Health, Boston, USA
5. University of Pittsburgh Graduate School of Public Health, Pittsburgh, PA, USA
6. University of Toronto, Toronto ON, Canada
7. The Hospital for Sick Children, Toronto ON, Canada

Please cite our publications when using this code:
 
- Jalal H, Pechlivanoglou P, Krijkamp E, Alarid-Escudero F, Enns E, Hunink MG. 
An Overview of R in Health Decision Sciences. Med Decis Making. 2017; 37(3): 735-746. 
https://journals.sagepub.com/doi/abs/10.1177/0272989X16686559
 
- Krijkamp EM, Alarid-Escudero F, Enns EA, Jalal HJ, Hunink MGM, Pechlivanoglou P. 
Microsimulation modeling for health decision sciences using R: A tutorial. 
Med Decis Making. 2018;38(3):400–22. 
https://journals.sagepub.com/doi/abs/10.1177/0272989X18754513
 
- Krijkamp EM, Alarid-Escudero F, Enns E, Pechlivanoglou P, Hunink MM, Jalal H. 
A Multidimensional Array Representation of State-Transition Model Dynamics. 
Med Decis Making. Online First https://doi.org/10.1177/0272989X19893973

Copyright 2017, THE HOSPITAL FOR SICK CHILDREN AND THE COLLABORATING INSTITUTIONS. 
All rights reserved in Canada, the United States and worldwide. Copyright, 
trademarks, trade names and any and all associated intellectual property are 
exclusively owned by THE HOSPITAL FOR Sick CHILDREN and the collaborating 
institutions. These materials may be used, reproduced, modified, distributed 
and adapted with proper attribution.


# 01 Load packages

```{r, warning = F, message = F}
if (!require('pacman')) install.packages('pacman'); library(pacman) # use this package to conveniently install other packages
# load (install if required) packages from CRAN
p_load("here", "dplyr", "devtools", "scales", "ellipse", "ggplot2", "lazyeval", "igraph", "truncnorm", "ggraph", "reshape2", "knitr", "stringr", "diagram", "bcea", "readxl")                                               
# load (install if required) packages from GitHub
# install_github("DARTH-git/dampack", force = TRUE) Uncomment if there is a newer version
# install_github("DARTH-git/dectree", force = TRUE) Uncomment if there is a newer version
p_load_gh("DARTH-git/dampack", "DARTH-git/dectree")
```

# 02 Load functions

```{r}
source('functions_WXYZ.R')
source("ImportVars.R")
```

# 03 Input model parameters

```{r}
# Strategy names
v_names_str <- c("Standard of Care", "New Treatment") 

# Number of strategies
n_str <- length(v_names_str)

# Markov model parameters
n_sim   <- 1000       # number of probabilistic simulations (arbitrary for deterministic model)
n_t     <- 20            # time horizon, number of cycles
v_n     <- c("W", "X", "Ytransition", "Y", "Ztransition", "Z")  # health state names  
n_states     <- length(v_n)         # number of health states

# Discounting factor
d_e  <- d_c <- 0.015

# calculate discount weights for costs for each cycle based on discount rate d_c
v_dwc <- 1 / (1 + d_e) ^ (0:n_t) 
# calculate discount weights for effectiveness for each cycle based on discount rate d_e
v_dwe <- 1 / (1 + d_c) ^ (0:n_t) 
```

```{r}
# Read in parameters using Batch Importer
Inputs <- read_excel("Model Inputs.xls") #The Excel file name
Inputs <- subset(Inputs, Value >-1) #Remove blank rows
varlist <- ImportVars(Inputs, n_sim)

# Specify a dataframe of parameter values
df_values <- data.frame(varlist$varmean)
colnames(df_values) <- varlist$varname
```

# 04 Define and initialize matrices and vectors

```{r}
# Assign parameter values to a list using 'get_values'
list_values <- get_values(df_values)

# Generate the transition matrix for 'notrt' and 'trt' using 'make_matrix'
list_tmatrix <- make_matrix(list_values)

# create the markov trace matrix M capturing the proportion of the cohort in each state 
# at each cycle
m_M_notrt <- m_M_trt <- matrix(NA, 
                                nrow     = n_t + 1, ncol = n_states,
                                dimnames = list(paste("cycle", 0:n_t, sep = " "), v_n))

# Define starting values for the first cycle
m_M_notrt[1, ] <- m_M_trt[1, ] <- c(list_values$p_W, list_values$p_X, 0, 0, 0, 0) 

# create the transition probability matrix for NO treatment
m_P_notrt  <- matrix(0,
                     nrow = n_states,
                     ncol = n_states,
                     dimnames = list(v_n, v_n)) # name the columns and rows of the matrix
```

# 05 Run Markov model

```{r}
for (t in 1:n_t){     # loop through the number of cycles
  # estimate the Markov trace for the next cycle (t + 1)
  m_M_notrt[t + 1, ]  <- t(m_M_notrt[t, ]) %*% list_tmatrix$m_P_notrt
  m_M_trt[t + 1, ]    <- t(m_M_trt[t, ])  %*% list_tmatrix$m_P_trt      
} # close the loop

```

# 06 Compute Cost-Effectiveness Outcomes

```{r}
# Vectors with costs and utilities by treatment
df_cea <- make_cea(list_values, m_M_notrt, m_M_trt, v_dwe, v_dwc)
# Print outputs using 'dampack'
calculate_icers(df_cea$Cost, df_cea$Effect, df_cea$Strategy)
```

